---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mStats

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mStats)](https://CRAN.R-project.org/package=mStats)
[![R-CMD-check](https://github.com/myominnoo/mStats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/myominnoo/mStats/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/myominnoo/mStats/branch/main/graph/badge.svg)](https://app.codecov.io/gh/myominnoo/mStats?branch=main)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

`mStats` provides an opinionated set of R wrapper functions for data analysis and manipulation, covering a range of tasks from data exploration and transformation to regression modeling and statistical summaries. It serves as my personal collection of functions developed based on my own preferences and experiences, tailored to meet my specific needs in various projects.

As an evolving package, I continuously add new functions to `mStats` to adapt to my evolving needs and preferences. This highly opinionated package reflects my unique perspective and incorporates the solutions I have found valuable in my data analysis workflows.

By using `mStats`, users can benefit from my expertise and the convenience of having a centralized set of functions designed to address common data manipulation, analysis, and visualization tasks. This personalized approach ensures efficiency and consistency across projects, as the package encompasses the techniques I prefer.

I believe that `mStats` makes a valuable contribution to the R community by offering insights and potential solutions derived from my own experiences. It showcases the power of customization and demonstrates how individual preferences can shape the development of tools to enhance data analysis workflows.

## Installation

Install from CRAN with :

```r
install.packages("mStats")
```

Or install the development version from GitHub :

``` r
# install.packages("devtools")
devtools::install_github("myominnoo/mStats")
```

## Usage

Below is a quick demonstration of what mStats can do.

```{r example}
library(mStats)
```

### codebook: check an overview of the data

```{r}
codebook(mtcars)
```

### cut 

